---
title: "Team"
author: "John Banta, Kelsey Mclnturff"
date: "10/31/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Github info

Owner username: jdbanta

Partner username: kmcintur

Repo: [ds202_lab5](https://github.com/jdbanta/ds202_lab5)

## Sections {.tabset}

### Processing the data

#### 1

```{r}
data<-read.table('diabetes.txt',header=TRUE)
head(data)
```

#### 2

```{r}
data[data==""]<-NA
data<-droplevels(data, "")
head(data)
```

#### 3

```{r}
vars<-names(data) %in% c("id", "bp.2s", "bp.2d")
diabetes_reduced<-data[!vars]
head(diabetes_reduced)
```

##### 4

```{r}
index.na=apply(is.na(diabetes_reduced), 1, any)
diabetes_clean<-diabetes_reduced[!index.na,]
```

#### 5

```{r}
dim(diabetes_clean)
```

This indicates that the dimensions of the new diabetes_clean dataset has 366 rows and 16 columns

### Exploring and transforming data

#### 6

#### 7

#### 8

### Visualizations

#### 10

#### 11

#### 12

### Messy data

#### 13

#### 14

### Regression models

#### 15

```{r}
fit = lm(glyhb ~stab.glu + age + waist + ratio+ factor(frame),data=diabetes_clean)
 summary(fit)
```

Looking first at the summary output, the F-statistic is 77.49 with an associated p-value of <.001 indicates that this model is useful for predicting the Glycosolated Hemoglobin with the combination of the input variables. Furthermore the adjusted R-squared is .557, which indicates that the model, after being adjusted for its complexity can still explain 55.7% of the variability in the response. However, it appears that the final two variables are not significant predictors of the Glycosolated Hemoglobin, so we could probably remove them to help simplify the model and avoid overfitting. 

#### 16

The estimate of stab.glu indicates that for every unit increase in stab.glu, with all other variables held constant, then the Glycosolated Hemoglobin will increase by .0035182. Similarly with the age, waist and ratio variables with increases of .0033632, .0047925, and .0219341 respectively. With the factor variables an individual with a medium frame is predicted to have .0309167 more Glycosolated Hemoglobin than the baseline of the large frame. Similarly with the small frame the model predicts .0131840 more than the baseline of the large frame. 

#### 17

```{r}
fit$fitted.values
```

These fitted values are predictions of Glycosolated Hemoglobin using the information of the given rows to predict it.

```{r}
pred<-data.frame(stab.glu = 90, age = 35, waist = 30, ratio = 5.1,frame = "small")
predict(fit,pred)
```

For someone with a stab.glu=90, age=35, waist=30, ratio=5.1 and a small frame the model predicts a value of 4.870237

#### 18

Inference is using a model to learn about the model generated by your data. However prediction is taking what you learned from the data and use it to predict for a datapoint that isn't in the dataset. 

#### 19

Linear regression will be only impacted by the inputs that go into the equation, however the knn regression model can be impacted by the amount of neighbors that you look at. The nice thing about knn regression is that it will be put together with data that are very similar, however with linear regression one variable change could drastically change the output.

### Reflection

#### 20